---
title: "Project_2A - Report Card"
author: "Fernando Figueres"
date: "March 9, 2019"
output: html_document
---

The dataset by proposed by Sheryl Piechocki, the details can be found [here](https://bbhosted.cuny.edu/webapps/discussionboard/do/message?action=list_messages&course_id=_1705327_1&nav=discussion_board&conf_id=_1845526_1&forum_id=_1891187_1&message_id=_31271363_1).


Based on the data we've analyzed the student has maintained an overall grade above 90% with her best performance on ACT Prep matreial which has an overall of 94.34%. This is spite of the poor performance during chapter 1 in which she scored 60% which was her lowest grade so far. It is encouraging to see that she improved her overall grade for the second trimester by almost 7%.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library imports}
suppressMessages(library(tidyverse))
```

```{r Tidy}
repcard <- 
  read_csv("repcard.csv", #import the csv file into a df
           col_types = 
             cols(
              `ACT Prep` = "i",
              Chapter = "i", 
              Classwork = "i", 
              Writing = "i"), 
           skip = 1) %>%
  as_tibble() %>% # Convert to a tibble
  mutate(Trimester = ifelse(Chapter < 5,1,2)) %>% #Add trimester as a variable
  rename(ACT_prep = `ACT Prep`) %>% #rename variable
  filter(!is.na(Chapter)) %>% #Remove NA values
  select(Trimester, everything()) %>% #Set Trimester as first variable
  gather("Component","Grade",3:5) #Set grade components as rows
```

```{r Chart Grade by Chapter  Component}
p <- ggplot(repcard, aes(x = Component, y = Grade, fill = Component))
p <- p + ggtitle("Grade by Chapter & Component")
p <- p + theme(legend.position = "bottom")
p <- p + geom_bar(stat = "identity", width = 0.95, position = "stack")
p <- p + facet_grid(. ~ Chapter)
p <- p + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
p
```

```{r}
p <- ggplot(data = repcard, aes(x = Chapter, y = Grade, fill = Component))
p <- p + ggtitle("Grade by Chapter & Component")
#p <- p + scale_y_continuous(labels = scales::percent)
p <- p + theme(legend.position = "none")
#p <- p + scale_fill_manual(values=c('#01426A','#8D0013'))
p <- p + geom_bar(stat = "identity", width = 0.95)
p <- p + facet_grid(. ~ Component)
p
```

```{r Grades by Component}
grades_by_c <- repcard %>%
  group_by(Component) %>%
  summarise(avg_grade = mean(Grade))

(grades_by_c)
```


```{r Grades by Chapter}
grades_by_chp <- repcard %>%
  group_by(Chapter) %>%
  summarise(avg_grade = mean(Grade))

(grades_by_chp)
```

```{r Grades by Trimester}
grades_by_tri <- repcard %>%
  group_by(Trimester) %>%
  summarise(avg_grade = mean(Grade))

(grades_by_tri)
```