---
title: "Project_2B - New York Weather"
author: "Fernando Figueres"
date: "March 9, 2019"
output: html_document
---

The dataset by proposed by Jimmy Ng, the details can be found [here](https://bbhosted.cuny.edu/webapps/discussionboard/do/message?action=list_messages&course_id=_1705327_1&nav=discussion_board&conf_id=_1845526_1&forum_id=_1891187_1&message_id=_31262338_1).

For simplicity, let's just focus on temperature and individual months. "Month" as a categorical variable is broken into twelve columns, and majority of cells in the table contains both Fahrenheit and Celsius values.  This is a "wide" structure, but we can transform it with tidyr::gather() to make it "long". Once we "gather" the columns, we can extract and split the values into Fahrenheit and Celsius by using the tidyr::separate(). Subsequently, we can "gather" the two columns (i.e. Fahrenheit, Celsius) into one and call it "temperature_scale". 

We can perform some EDA on this data frame, such as what is the average hottest/coldest month in degree Celsius, and then we can use ggplot2 to create simple visualization to aid our analysis, e.g. barplot or dotplot to sort and line up the month by temperature. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressMessages(library(xml2))
suppressMessages(library(rvest))
suppressMessages(library(tidyverse))
```


```{r}
df <- 
  read_html("https://en.wikipedia.org/wiki/Climate_of_New_York") %>%         
  html_nodes("table") %>%
    .[[3]] %>%        
  html_table(trim = TRUE)
```

```{r}
df2 <- df[1:15,] %>% 
  setNames(as.character(df[1,])) %>% 
  rename(Measure = Month) %>% 
  filter(str_detect(Measure,"(°C)")) %>% 
  gather("Month","Temp",2:13) %>% 
  separate(Temp, into = c("F", "C"),sep = "\\(") %>% 
  mutate(C = str_replace(C, "\\)", "")) %>% 
  mutate(Measure = str_replace(Measure, "°F \\(°C\\)", "")) %>% 
  gather("Scale","Temp",4:5) %>% 
  mutate(Temp = type.convert(Temp)) %>% 
  select (-c(Year))

head(df2)
```

 