---
title: "Project_2C - Median Rental Price/sqft by State"
author: "Fernando Figueres"
date: "March 9, 2019"
output: html_document
---

The dataset by proposed by Sunny Mehta, the details can be found [here](https://bbhosted.cuny.edu/webapps/discussionboard/do/message?action=list_messages&course_id=_1705327_1&nav=discussion_board&conf_id=_1845526_1&forum_id=_1891187_1&message_id=_31286554_1).

From the histogram, we can see that the overall distribution of rental prices is right skewed. if we then look and the ranked barplot, we see that a few locations have much higher rent prices than the rest of the country.

If we look at the yearly changes we find that New York experienced the sharpest drop in prices (-17.5%) and Wyoming increased the most (8.8%).

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Import and Tidying

```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(lubridate))
suppressMessages(library(RMySQL))
```


```{r Connect to Database}
mdb <- DBI::dbConnect(RMySQL::MySQL(),
  dbname='project2db',
  host = "35.224.164.217",
  user = "root",
  password = "2Password"#rstudioapi::askForPassword("Database password")
)

#code borrowed from https://dbplyr.tidyverse.org/articles/dbplyr.html
```


##Define SQL query to extract required data
```{r}
sql <- "SELECT * FROM project2db.RentPrice"
```


##Data is retreived and loaded into a data frame
```{r}
medrent <- 
  dbGetQuery(mdb, sql) %>% 
  gather(Month,Rent,3:14) %>% #Gather months into rows
  mutate(Month = as.Date(paste(Month,"-01",sep=""))) %>% #Convert strings to dates 
  mutate(Month = month(Month,label = TRUE, abbr = FALSE)) #Convert dates to month

head(medrent)
```

##Find the current (latest) distribution.

```{r}
medrent12 <- #Create a new df that includes only the latest month
  medrent %>% 
  filter(Month == "December")
```

```{r}
ggplot(medrent12, aes(Rent, fill = Rent)) +
  ggtitle("Distribution of Median Rental Price/sqft by State (Dec 2018, Zillow)")+
  geom_histogram(binwidth = 0.25, fill="steelblue") +
  theme(legend.position = "none")
```

```{r fig.width=7, fig.height=8}
p<-ggplot(data=medrent12, aes(x=reorder(RegionName,Rent), y=Rent), width=.5,position = position_dodge(width = 60)) +
  ggtitle("Median Rental Price/sqft by State (Dec 2018, Zillow)")+
  geom_bar(stat="identity", fill="steelblue")+
  theme_minimal() +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  coord_flip() 
p
```

## Yearly % Change by State

```{r}
medrent2 <- 
  read_csv("State_MedianRentalPricePerSqft_AllHomes.csv", #Impor the CSV file
    col_types = cols(SizeRank = col_integer())) %>% 
  mutate(ychange = ((`2018-12`/`2018-01`)-1)) %>% 
  select (-c(3:14))

head(medrent2)
```

```{r fig.width=7, fig.height=8}
p<-ggplot(data=medrent2, aes(x=reorder(RegionName,ychange), y=ychange), width=.5) +
  ggtitle("% Chg in Median Rental Price/sqft by State (2018, Zillow)")+
  geom_bar(stat="identity", fill="steelblue")+
  theme_minimal() +
  #coord_fixed(ratio = 1, xlim = NULL, ylim = NULL) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent)
p
```



 